---
title: "Hw7_Ex_4"
author: "Manuel Alejandro Garcia Acosta"
date: "10/30/2019"
output: pdf_document
---

```{r}
library(BioStatR)
library(olsrr)
```

# Exercise 4 Homework 7

## Read the data

```{r}
setwd('/home/noble_mannu/Documents/PhD/First/STAT_2131_Applied_Statistical_Methods_I/HW7')
steam_data <- read.table('steam_text.txt', header = TRUE)
```

## Making the regression model

```{r}
m1 <- lm(steam ~ fat+glycerine+wind+frezday+temp, data = steam_data)
```

We display the summary of our model.

```{r}
summary(m1)
```

# Performing forward and backward model selection

## Performing forward selection

```{r}
alpha.1 <- 0.1
forward.1 <- ols_step_forward_p(m1, penter = alpha.1)
forward.1$predictors   # Model only contains temp and fat
# Our final model only has temp and fat as predictors
```

## Performing backward selection

```{r}
alpha.2 <- 0.2
backward.1 <- ols_step_backward_p(m1, penter = alpha.2)
backward.1$removed    # Model only contains temp and fat
# Our final model only has temp and fat as predictors
```

Although we weren't required to do so, I performed forward step-wise selection just to see how it went.

### Step-wise forward

```{r}
step.wise <- ols_step_both_p(m1, pent = alpha.1, prem = alpha.2)
step.wise$predictors  # Model only contains temp and fat
# Once again our final model only has temp and fat as predictors
```

Now we'll run best subset selection for this model

# Running best subset selection

```{r}
Best.subset.steam <- olsrr::ols_step_best_subset(m1)
```

## Choosing the model based on AIC

```{r}
which.min(Best.subset.steam$aic)
# Returns row 2, this corresponds to the model with only fat and temp as predictors
# Prints the names of the predictors used in the best model with AIC criteria
Best.subset.steam$predictors[which.min(Best.subset.steam$aic)]
# This shows that the best model using AIC criteria is the one that has fat and temp as predictors
```


## Choosing the model based on BIC

```{r}
which.min(Best.subset.steam$sbc)
# Returns row 2, this corresponds to the model with only fat and temp as predictors
# Prints the names of the predictors used in the best model with BIC criteria
Best.subset.steam$predictors[which.min(Best.subset.steam$sbc)]
# This shows that the best model using BIC criteria is the one that has fat and temp as predictors
```

We conclude that backward, forward, step-wise forward and AIC and BIC methods all returned back the model with only 'fat' and 'temp' as predictors.

### Note: With adjusted R2 criteria we get the same model as above

```{r}
which.max(Best.subset.steam$adjr)
# Returns row 2, this corresponds to the model with only fat and temp as predictors
# Prints the names of the predictors used in the best model with adjusted R2 criteria
Best.subset.steam$predictors[which.max(Best.subset.steam$adjr)]
```

### Note: With R2 criteria we get the full model

```{r}
which.max(Best.subset.steam$rsquare)
# Returns row 5, this corresponds to the full model
# Prints the names of the predictors used in the best model with R2 criteria
Best.subset.steam$predictors[which.max(Best.subset.steam$rsquare)]
```