---
title: "Model Selection"
author: "Chris McKennan"
date: "10/24/2019"
output: html_document
---

Here we will do model selection based on the Boston housing price data. Variables we will consider are:
mvalue = median value of owner-occupied homes in \$1000. This is our Y.
crim = per capita crime rate by town
zn = proportion of residential land zoned for lots over 25,000 sq.ft
indus = proportion of non-retail business acres per town
chas = Charles River factor variable (= 1 if tract bounds river; 0 otherwise)
nox = nitric oxide concentration (parts per 10 million)
rooms = average number of rooms per dwelling
age = proportion of owner-occupied units built prior to 1940
distance = weighted distances to five Boston employment centers
radial = index of accessibility to highways
tax = full-value property-tax rate per \$10,000
pt = pupil-teacher ratio by town


#Load data
```{r LoadData}

setwd('/home/noble_mannu/Documents/PhD/First/STAT_2131_Applied_Statistical_Methods_I/HW7')
Data <- data.frame(read.table(file = "Boston.txt", header = T, sep = "\t", stringsAsFactors = F))
```

#Function to create histogram scatterplots
```{r HistScatter}
#install.packages("BioStatR")
library(BioStatR)
panel.hist <- function(x, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}
```

#Explore the data
```{r ExploreData}
head(Data)
pairs(~mvalue + indus + chas + nox + rooms + age + distance + radial + tax + pt, data=Data,
   main="Scatterplot Matrix")

pairs(~mvalue + crim + nox + rooms + distance + tax + pt, data=Data,
   main="Scatterplot Matrix", diag.panel = panel.hist)
```

#Full model
```{r FullModel}
full.fit <- lm(mvalue ~ ., data=Data)

summary(full.fit)
plot(predict(full.fit), residuals(full.fit))
qqnorm(residuals(full.fit)); qqline(residuals(full.fit), col="red")
```

#Best subset variable selection
Look at ALL 2^11 = 2,048 possible models and choose the best one.
```{r BestSubset}
#install.packages("olsrr")

Best.subset <- olsrr::ols_step_best_subset(full.fit)   #Run best subset selection. This may take a minute or two...
```

##R2 and adjusted R2
Choose the model based on R2 or adjusted R2
```{r R2andAdjustedR2}
which.max(Best.subset$rsquare)
which.max(Best.subset$adjr)

predictors.include <- strsplit(Best.subset$predictors[which.max(Best.subset$adjr)], "[ ]+", perl=T)[[1]]
colnames(Data)[!(colnames(Data)%in%predictors.include)]  #Which is left out?
```
Based on adjusted R2, we include all predictors EXCEPT for indus = proportion of non-retail business acres per town

##AIC
```{r AIC}
which.min(Best.subset$aic)
```
If we know that the best AIC-determined model has 10 predictors, why do we know it is the same as above?

##BIC
Called SBC in output of olsrr::ols_step_best_subset
```{r BIC}
#Let's calculate BIC from AIC#
AIC <- Best.subset$aic
our.BIC <- AIC - 2*(1:11) + log(nrow(Data))*(1:11)

#How does this compare to their BIC#
Best.subset$sbc - our.BIC   #We differ by a constant. Why does this not matter?

which.min(our.BIC)
```

#Forward model selection

```{r ForwardModelSelection}
alpha.1 <- 0.2
forward.1 <- ols_step_forward_p(full.fit, penter = alpha.1)
forward.1$predictors   #Contains all 11 predictors
p.values.fullmodel <- summary(full.fit)$coefficients[,4]   #Why does model contain all 11 predictors when indus p-value from the full model fit is > alpha.1?

#Fit model when indus enters#
indus.enter <- lm(mvalue~ rooms + pt + nox + distance + crim + chas + age + zn + indus, data=Data)
summary(indus.enter)

#What if we change alpha.1?
alpha.1 <- 0.01   #Harder for covariate to enter model
forward.2 <- ols_step_forward_p(full.fit, penter = alpha.1)
forward.2$predictors
```

#Backwards model selection

```{r BackwardsModelSelection}
alpha.2 <- 0.2
backward.1 <- ols_step_backward_p(full.fit, penter = alpha.2)
backward.1$removed    #Model contains all covariates, except for indus.
```

#Forward step-wise selection

```{r ForwardStepWiseSelection}
alpha.1 <- 0.1
alpha.2 <- 0.3
step.wise <- ols_step_both_p(full.fit, pent = alpha.1, prem = alpha.2)

step.wise$predictors  #No indus, exactly as we would expect
```